Secure Composition of Robust and Optimising Compilers

### 2023-07-05 popl/main.tex
Pass over BG

Programs ($\runtimetermvar$) without free variables are considered {\em whole}.
Contrary, they are {\em partial} if they contain variables without an accompanying definition.
Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.\MK{define final?}
Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
All multi-step relations filter $\emptyevent$ events.

The {\em behavior} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is a set containing all traces the program may emit during execution.
%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
The behavior of programs gives rise to the definition of satisfaction:

%\begin{definition}[Property Satisfaction]\label{def:propsat}
%  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\lift{\pi}$.
%\end{definition}
\begin{definition}[Class Satisfaction]\label{def:classsat}
  A whole program $\runtimetermvar$ satisfies $\class$ ($\sat{\runtimetermvar}{\class}$) iff $\behav{\runtimetermvar}\in\class$.
\end{definition}

\subsection{Attacker Model and Robust Trace-Property Preservation}\label{subsec:bg:rtp}

Partial programs cannot be executed, because they lack definitions for some symbols.
Providing definitions for symbols is usually the job of the \emph{linker}.
Notationally, linking two programs $\progvar[_{1}],\progvar[_{2}]$ is written $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
For sake of simplicity, linking does not fail.
With linking, it is possible to define robust satisfaction analogously to \Cref{def:propsat,def:classsat}:

\begin{definition}[Robust Property Satisfaction]\label{def:proprsat}
  A program $\progvar$ robustly satisfies $\pi$ ($\rsat{\progvar}{\pi}$) iff for any program $\contextvar$ it holds that $\sat{\link{\contextvar}{\progvar}}{\pi}$. (similarily for $\rsat{\progvar}{\class}$)
\end{definition}

Robust satisfaction gives rise to a maximally powerful attacker model: Any other partial program, here referred to as context $\contextvar$, is the attacker.
So, for example, if the programming language of $\progvar$ and $\contextvar$ allows guessing of memory addresses, then $\contextvar$ is able to read and write any position in memory and therefore guess the location of secrets.

It is expected that compilers preserve robust satisfaction.
In this paper, compilers are modelled as function from one programming language's syntax to the other.
Notationally, $\cc{\src{L}}{\trg{L}}$ denotes a compiler from $\src{L}$ to $\trg{L}$ and the paper refers to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
Now, define the compiler-property robust trace-property preservation.

\begin{definition}[Robust Trace-Property Preservation]\label{def:rtp}
  A compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of properties $\class$ ($\rtp{\cc{\src{L}}{\trg{L}}}{\class}$) iff for any property $\pi\in\class$ and $\src{p}$ written in $\src{L}$, given $\rsat{\src{\progvar}}{\pi}$, then $\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$.
\end{definition}

Note that in the conclusion of \Cref{def:rtp} there is a target-level context.
So, programs that robustly satisfy a property are hardened against a maximally malicious environment with respect to the power of the target language.

\section{Composing Properties\pages{3}}\label{sec:compprop}
\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that the case study (\Cref{sec:casestud}) uses.
By means of taking these properties as an example, \Cref{subsec:monitors} defines monitors for them and an agreement relation between a monitor's trace model with the most general trace model.
Then, monitor-based satisfaction is introduced and it is proven that trace-property satisfaction is implied by it.
Since monitor-based satisfaction is easier to reason about, this simplifies later proofs considerably.

>>>>

A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
Programs without free variables are considered {\em whole}.
Contrary, they are {\em partial} if they contain variables without an accompanying definition.
Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.
Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
All multi-step relations filter $\emptyevent$ events.

The {\em behavior} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is a set containing all traces the program may emit during execution.
%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
Finally, the satisfaction of a class (or lifted property) is defined as follows:

\begin{definition}[Satisfaction]\label{def:propsat}
  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\pi$.
\end{definition}

As described earlier, partial programs cannot be executed, because they lack definitions and, thus, get stuck.
{\em Linking} ($\link{\progvar[_{1}]}{\progvar[_{2}]}$) syntactic descriptions $\progvar[_{1}],\progvar[_{2}]$ can give rise to a whole program.
Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.
However, for convenience, linking of programs $\runtimetermvar[_{1}],\runtimetermvar[_{2}]$ shall be defined if both $\runtimetermvar[_{1}]$ and $\runtimetermvar[_{2}]$ have an initial dynamic state.
As typically done in a secure compilation context~\cite{abate2019jour}, this paper also assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$), which are just any syntactic description.
The component part of a whole program is the main object of interest and thus a known syntactic description.
In contrast, the context is usually arbitrary and thus has the role of an active {\em attacker} that can interact with the component by means of whatever features the programming language has.
The following definition extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:

\begin{definition}[Robust Satisfaction]\label{def:proprsat}
  A program $\runtimetermvar$ robustly satisfies $\pi$ ($\rsat{\runtimetermvar}{\pi}$) iff for any partial program $\contextvar$ it holds that $\sat{\link{\contextvar}{\runtimetermvar}}{\pi}$.
\end{definition}

\subsection{Secure Compilers}\label{subsec:bg:rtp}

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
The syntactic description of programs written in $\src{L}$ is denoted as $\src{\progvar}$, while for ones written in $\trg{L}$ this paper writes $\trg{\progvar}$.
It is considered good if this translation is semantics-preserving, i.e., correct.
Correct compilers may still be insecure compilers.
Programs translated by insecure compilers may violate security properties that were assumed to hold, given the presence of a context--level program.
So, ideally, a compiler is secure~\cite{abate2019jour} in the sense of the following definition:

\begin{definition}[Robust Preservation]\label{def:rtp}
  A compiler $\cc{\src{L}}{\trg{L}}$ robustly--preserves a class of properties $\class$ ($\rtp{\cc{\src{L}}{\trg{L}}}{\class}$) iff for any property $\pi\in\class$ and component $\src{p}$ such that $\rsat{\src{\progvar}}{\pi}$, it holds that $\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$.
\end{definition}

\section{Example Properties and their Monitors\pages{3}}\label{sec:compprop}
\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that the case study (\Cref{sec:casestud}) uses.
By means of taking these properties as an example, \Cref{subsec:monitors} defines monitors for them and an agreement relation between a monitor's trace model with the most general trace model.
Then, monitor-based satisfaction is introduced and it is proven that trace-property satisfaction is implied by it.
Since monitor-based satisfaction is easier to reason about, this simplifies later proofs considerably.
